RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-10-23 formalframework.tex
finished all algorithms

\begin{definition}[Attack Link]
\label{def:link:attack}
Given a set of arguments $Arg$, an attack link $att\in Arg\times Arg$ is a relation such that $att=(A_i,A_j)$ means:
\begin{itemize}
\item $A_i\in Arg$ is the unique identifier of the element that is performing that attack,
\item $A_j\in Arg$ is the unique identifier of the element that is being attacked.
\end{itemize}
Intuitively, $att=(A_i,A_j)$ means that argument $A_i$ is attacking argument $A_j$. A set of attack links is denoted by $Att$.
\end{definition}

@@

Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models. Both of these two translation algorithms are straightforward. Their main task is to ensure that in RationalGRL models, the GRL elements and links are wrapped in an argument with the corresponding type.

@@

\begin{algorithm}[h]
  \caption{GRL to RationalGRL Translation}
  \label{alg:translation:to-rationalgrl}
  \begin{algorithmic}[1]
    \Procedure{$TranslateToRationalGRL$}{$GRL$}
    \State $Arg \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:init}
    \For{$ie\in IE$}\label{alg:translation:to-rationalgrl:ie}
      \State $Arg \leftarrow Arg \cup \{(ie,ie.type)\}$
    \EndFor
    \For{$act\in Act$}
      \State $Arg \leftarrow Arg \cup \{(act, Actor)\}$
    \EndFor
    \For{$actIE\in ActIE$}
      \State $Arg \leftarrow Arg \cup \{(actIE, ActIE)\}$
    \EndFor
    \For{$contr\in Contr$}
      \State $Arg \leftarrow Arg \cup \{(contr, Contr)\}$
    \EndFor
    \For{$dep\in Dep$}
      \State $Arg \leftarrow Arg \cup \{(dep, Dep)\}$
    \EndFor
    \For{$decomp\in Decomp$}
      \State $Arg \leftarrow Arg \cup \{(decomp, Decomp)\}$
    \EndFor
    \State \Return $(Arg, \emptyset)$\label{alg:translation:to-rationalgrl:return}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption{RationalGRL to GRL Translation}
  \label{alg:translation:to-grl}
  \begin{algorithmic}[1]
    \Procedure{$TranslateToGRL$}{$RatGRL$}
    \State $E \leftarrow ComputeExtension(Arg,Att)$
    \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Contr\leftarrow\emptyset$
    \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
    \For{$(Elem, T)\in E$}\label{alg:translation:to-grl:extension-iter}
      \Switch{$T$}\label{alg:translation:to-grl:switch}
          \Case{$T\in Type$}
            \State $IE\leftarrow IE\cup \{Elem\}$
          \EndCase
          \Case{$T = Actor$}
            \State $Act\leftarrow Act \cup\{Elem\}$
          \EndCase
          \Case{$T=Contr$}
            \State $Contr\leftarrow Contr\cup \{Elem\}$
          \EndCase
          \Case{$T=Dep$}
            \State $Dep\leftarrow Dep\cup \{Elem\}$
          \EndCase
          \Case{$T=Decomp$}
            \State $Decomp\leftarrow Decomp\cup\{Elem\}$
          \EndCase
      \EndSwitch
    \EndFor
    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$\label{alg:translation:to-grl:return}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

@@

In the previous section we formalized a \emph{static} representation of the RationalGRL framework. In this section we formalize the \emph{dynamics}. We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model (Def.~\ref{def:rationalgrl-model}). These algorithms produce new elements and attack relations. We can then use argumentation semantics (Def.~\ref{def:semantics}) to compute sets of accepted arguments. The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.

\paragraph{Argument Schemes}

In all of the following algorithms, we assume the following
\begin{itemize}
\item There algorithms are being applied to some valid RationalGRL model $RatGRL$,
\item The procedure $mindId()$ generates a new unique id.
\end{itemize}

The algorithms for applying argument schemes are short, and consist simply of adding an argument for the element that is being added.

\begin{algorithm}[h]
  \caption{AS0: Add a new actor with name $n$}\label{alg:as0}
  \begin{algorithmic}[1]
    \Procedure{$AS_0$}{$n$}
    \State $actor \leftarrow (mintId(), n)$ \label{alg:as0:1}
    \State $Arg\leftarrow Arg \cup \{(actor,Actor)\}$\label{alg:as0:2}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} The algorithm takes one argument, namely the name of the actor $n$. On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$. On line~\ref{alg:as0:2} the pair $(actor,Actor)$ is added as a new argument, indicating that an argument of type $Actor$ is added.

\begin{algorithm}[h]
  \caption{Applying AS1: Actor with id $i$ has resource $n$}\label{alg:as1}
  \begin{algorithmic}[1]
    \Procedure{$AS_1$}{$i, n$}
    \State $rId\gets mintId()$
    \State $r\leftarrow (rId, Resource, n, AND)$
    \State $Arg\gets Arg\cup \{(r, Resource),(i,rId)\}$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Explain here...

\begin{algorithm}[h]
  \caption{Applying AS5: Goal with id $i$ decomposes into tasks $t$}\label{alg:as5}
  \begin{algorithmic}[1]
    \Procedure{$AS_5$}{$i, t$}
    \State $tId\gets mintId()$
    \State $t\leftarrow (tId, Task, t, AND)$
    \State $Arg\gets Arg\cup \{(t, Task),(i,tId)\}$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Explain here.

The remaining argument schemes are all very similar, in the sense that they all simply add a single argument to the set of arguments containing the element to be added. We therefore have omitted them here.

\subsubsection*{Algorithms for critical questions}

>>>>

\begin{definition}[Attack Link]
\label{def:link:attack}
Given a set of arguments $Arg$, an attack link $att\in Arg\times Arg$ is a relation such that $att=(A_i,A_j)$ means:
\begin{itemize}
\item $A_i\in Arg$ is the argument performing that attack,
\item $A_j\in Arg$ is the argument being attacked.
\end{itemize}
Intuitively, $att=(A_i,A_j)$ means that argument $A_i$ is attacking argument $A_j$. A set of attack links is denoted by $Att$.
\end{definition}

@@

Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models. Both of these two translation algorithms are straightforward. Their main task is to ensure that in RationalGRL models, the GRL elements and links are wrapped in an argument with the corresponding type. For each algorithm, we provide an explanation an an example of a translation.

@@

\begin{algorithm}[h]
  \caption{GRL to RationalGRL Translation}
  \label{alg:translation:to-rationalgrl}
  \begin{algorithmic}[1]
    \Procedure{$ToRationalGRL$}{$GRL$}
    \State $Arg \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:init}
    \For{$ie\in IE$}\label{alg:translation:to-rationalgrl:ie}
      \State $Arg \leftarrow Arg \cup \{(ie,ie.type)\}$
    \EndFor
    \For{$act\in Act$}
      \State $Arg \leftarrow Arg \cup \{(act, Actor)\}$
    \EndFor
    \For{$actIE\in ActIE$}
      \State $Arg \leftarrow Arg \cup \{(actIE, ActIE)\}$
    \EndFor
    \For{$contr\in Contr$}
      \State $Arg \leftarrow Arg \cup \{(contr, Contr)\}$
    \EndFor
    \For{$dep\in Dep$}
      \State $Arg \leftarrow Arg \cup \{(dep, Dep)\}$
    \EndFor
    \For{$decomp\in Decomp$}
      \State $Arg \leftarrow Arg \cup \{(decomp, Decomp)\}$
    \EndFor
    \State \Return $(Arg, \emptyset)$\label{alg:translation:to-rationalgrl:return}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption{RationalGRL to GRL Translation}
  \label{alg:translation:to-grl}
  \begin{algorithmic}[1]
    \Procedure{$ToGRL$}{$RatGRL$}
    \State $E \leftarrow ComputeExtension(Arg,Att)$
    \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Contr\leftarrow\emptyset$
    \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
    \For{$(Elem, T)\in E$}\label{alg:translation:to-grl:extension-iter}
      \Switch{$T$}\label{alg:translation:to-grl:switch}
          \Case{$T\in Type$}
            \State $IE\leftarrow IE\cup \{Elem\}$
          \EndCase
          \Case{$T = Actor$}
            \State $Act\leftarrow Act \cup\{Elem\}$
          \EndCase
          \Case{$T=Contr$}
            \State $Contr\leftarrow Contr\cup \{Elem\}$
          \EndCase
          \Case{$T=Dep$}
            \State $Dep\leftarrow Dep\cup \{Elem\}$
          \EndCase
          \Case{$T=Decomp$}
            \State $Decomp\leftarrow Decomp\cup\{Elem\}$
          \EndCase
      \EndSwitch
    \EndFor
    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$\label{alg:translation:to-grl:return}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

@@

The following theorem follows directly from the translation algorithms.

\begin{theorem}[Translation correctness]
\label{th:translation-correctness}
Given some valid GRL model (Def.~\ref{def:valid-grl-model}, translating it to RationalGRL and then back to GRL results in the same valid GRL mode, i.e.
$$ToGRL(ToRationalGRL(GRL)) = GRL.$$
\end{theorem}

@@

In the previous subsection we formalized a \emph{static} representation of the RationalGRL framework. In this section we formalize the \emph{dynamics}. We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model (Def.~\ref{def:rationalgrl-model}). These algorithms are applied to RationalGRL models and produce new arguments and attack relations. We can then use argumentation semantics (Def.~\ref{def:semantics}) to compute sets of accepted arguments. The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.

As discussion in Section~\ref{sect:overview}, ll the argument schemes and critical questions of Table~\ref{table:argument-schemes} fall into one of the following four categories:
\begin{itemize}
\item \emph{INTRO}: These algorithms add one or multiple new arguments to the RationalGRL model, without creating any attack links. The arguments are for new GRL elements or links.
\item \emph{DISABLE}: These algorithms add a single argument to the RationalGRL model, which attacks one of multiple existing arguments. The counterargument that is added does not correspond to any GRL element, but instead disables an existing one by attacking all arguments for it.
\item \emph{REPLACE}: These algorithms can be seen as a combination of \emph{INTRO} and \emph{DISABLE}. They add a new argument corresponding to a GRL element or link, and this new argument attacks a previous version of the element or link. 
\end{itemize}

The three subsections of this section correspond to these three types of arguments.

In all of the following algorithms, we assume the following
\begin{itemize}
\item There algorithms are being applied to some valid RationalGRL model $RatGRL$ (Def.~\ref{def:valid-rationalgrl-model}),
\item The procedure $mindId()$ generates a new unique id.
\end{itemize}

\subsubsection{INTRO algorithms}

The following arguments schemes and critical questions of Table~\ref{table:argument-schemes} fall into this category:
\begin{itemize}
\item AS0-AS12
\item CQ5b, CQ6b, CQ6c, CQ7b, CQ9, CQ10b
\end{itemize}

These type of algorithms are short, and consist simply of adding an argument for the element that is being added. Since the structure of these algorithms are all very similar, we do not discuss them all.

\begin{algorithm}[h]
  \caption{AS0: $a$ is an actor}\label{alg:add-actor}
  \begin{algorithmic}[1]
    \Procedure{$AS_0$}{$n$}
    \State $actor \leftarrow (mintId(), a)$ \label{alg:as0:1}
    \State $Arg\leftarrow Arg \cup \{(actor,Actor)\}$\label{alg:as0:2}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Algorithm takes one argument, namely the name of the actor $a$. On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$. On line~\ref{alg:as0:2} the pair $(actor,Actor)$ is added as a new argument, indicating that an argument of type $Actor$ is added.

Consider for instance...\todo{Marc}{Marc}{Add example}

\begin{algorithm}[h]
  \caption{AS1: Actor with id $i$ has resource $R$}\label{alg:add-actor}
  \begin{algorithmic}[1]
    \Procedure{$AS_1$}{$i, R$}
    \State $res\_id\gets mintId()$\label{alg:add-actor:id}
    \State $res\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-actor:res}
    \State $Arg\gets Arg\cup \{(res, Resource),(i,res\_id)\}$\label{alg:add-actor:arg}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} The original argument schemes AS1 is phrased as ``Actor $a$ has resource $R$''. We have slightly reworded this in Algorithm~\ref{alg:add-actor} to ``Actor with id $i$ has resource $R$''. The reason is that in GRL, actors have identifiers, so since we assume actor $a$ exists already, we also assume an identifier for it exists. It would be straightforward to introduce a mapping from actor names to identifiers, but we have no done so here for simplicity.

The algorithm itself runs as follows: on line~\ref{alg:add-actor:id}, a unique id is assigned to variabel $res\_id$ (``resource identifier''). On line~\ref{alg:add-actor:res}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A$. Finally, on line~\ref{alg:add-actor:arg}, two arguments are added to the set of arguments $Arg$: one for the resource, and one for the containment relations between the actor and the resource (Def.~\ref{def:act-ie-relation}.

Since arguments schemes AS2-AS4 are very similar to AS1, we have omitted it here.

\begin{algorithm}[h]
  \caption{AS5: Goal with id $i$ decomposes into task $t$}\label{alg:add-decomp}
  \begin{algorithmic}[1]
    \Procedure{$AS_5$}{$i, t$}
    \State $task\_id\gets mintId()$\label{alg:add-decomp:task-id}
    \State $task\leftarrow (task\_id, Task, t, AND)$\label{alg:add-decomp:task}
    \State $Arg\gets Arg\cup \{(task, Task),((mintId(),i,task\_id), Decomp)\}$\label{alg:add-decomp:arg}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Similar to the previous algorithm, we have slightly reworded critical question AS5. We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $t$ is a decomposition of $G$.

In Algorithm~\ref{alg:add-decomp}, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:task}. Finally, on line~\ref{alg:add-decomp:arg}, two arguments are added to the set of arguments $Arg$. The first argument $(task,Task)$ is simply an argument for the task that is added. The second argument $((mintId(),i,task\_id),Decomp)$ is an argument for the decomposition link $(mintId(), i, task\_id)$ (Def.~\ref{def:link:decomp}, going from the existing goal with identifier $i$ to the new task with identifier $task\_id$.

The remaining argument schemes AS6-AS-12 are all very similar to the previous algorithms and have been omitted here.

The critical questions of type \emph{INTRO} are very similar as well, with one exception: the require an answer. For instance, suppose CQ5b: ``Does goal $G$ decompose into other tasks?'' is answered with: ``Yes, namely into task $T$''. In this case, we simply obtain an instantiation of argument scheme AS5: ``Goal $G$ decomposes into task $T$'', which can be executed with Algorithm~\ref{alg:add-decomp}. This is the same for all the other critical questions of type \emph{INTRO} as well. Therefore, we have omitted them here as well.

\subsubsection{DISABLE algorithms}

As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument, which is an argument for an existing GRL element or link. The argument that is added is itself not an argument for a GRL element or link.

In all of these algorithms, we assume the critical question is answered affirmatively. For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No'' (see right-most column of Table~\ref{table:argument-schemes}).

\begin{algorithm}[h]
  \caption{CQ0: Is actor with id $i$ relevant? No}\label{alg:actor-not-relevant}
  \begin{algorithmic}[1]
    \Procedure{$CQ_0$}{$i$}
    \State $A \leftarrow ((mintId(),CQ0),GenArg)$\label{alg:actor-not-relevant:genarg}
    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:actor-not-relevant:genarg2}
    \For{$A_i\in \{(actor_i,Actor)\in Arg\}$}\label{alg:actor-not-relevant:for}
      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:actor-not-relevant:att}
    \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered affirmatively (i.e., with ``No''). First, on lines~\ref{alg:actor-not-relevant:genarg} and~\ref{alg:actor-not-relevant:genarg2}, an argument is created for the critical question and added to the set of arguments $Arg$. Since this argument is not an argument for a GRL element or link, it is formalized as a \emph{generic counterargument} $((mintId(), CQ0), GenArg)$ (Def.~\ref{def:generic-argument}). The for loop starting at line~\ref{alg:actor-not-relevant:for} then iterates over ll arguments $(actor_i,Actor)$, where $i$ is the id of the actor that is no longer relevant. Recall from Def.~\ref{def:notation} that we denote with $actor_i$ an actor with identifier $i$. Thus, this for loop iterates over all arguments for actors with identifier $i$. The reason why there could be multiple of such actors is that the actor can be refined by an algorithm of type \emph{REPLACE}. We will explain this in more detail in the example below. Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ that is created to the argument for the actor $A_i$. After executing the algorithm, all existing arguments for the actor with identifier $i$ are attacked by a newly created argument $A$.

Consider for example...\todo{Marc}{Marc}{Add example here}

Critical questions C1-CQ3 are all very similar to CQ0 and have therefore been omitted here.

\begin{algorithm}[h]
  \caption{CQ5a: Does the goal with id $g\_id$ decompose into task with id $t\_id$? No}\label{alg:no-decomp}
  \begin{algorithmic}[1]
    \Procedure{$CQ5a$}{$g\_id,t\_id$}
    \State $A \leftarrow ((mintId(),CQ5a),GenArg)$\label{alg:no-decomp:genarg}
    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:no-decomp:genarg2}
    \For{$A_i\in \{((k,m,n),Decomp)\in Arg\mid $\\
    \qquad\qquad\qquad $m=g\_id, n=t\_id\}$}\label{alg:no-decomp:for}
      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:no-decomp:att}
    \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links. This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links $(k,m,n)$ (Def.~\ref{def:link:decomp}) such that $m$ equals the goal identifier $g\_id$, and $n$ equals the task identifier $t\_id$. This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att}.

Almost all of the remaining critical questions are similar in structure. CQ11 (``Is the element relevant/useful'') is slightly different since the attack element is not of a specific type, but is simply any GRL element. However, the resulting algorithm is very similar to the previous two and has therefore been omitted. The only algorithm of type \emph{DISABLE} that we still discuss is $Att$.

\begin{algorithm}[h]
  \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$}\label{alg:generic}
  \begin{algorithmic}[1]
    \Procedure{$Att$}{$A_1,\ldots,A_n$}
    \State $A \leftarrow ((mintId(),Att),GenArg)$\label{alg:generic:arg}
    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:generic:addarg}
    \For{$A_i\in \{A_1,\ldots,A_n\}$}\label{alg:generic:for}
      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:generic:att}
    \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counter-arguments to arguments. In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition. In this algorithm, however, \emph{any} set of arguments can be attacked by a new argument.

\todo{Marc}{Marc}{Give example}

\subsubsection{REPLACE algorithms}

Recall that the \emph{REPLACE} algorithms both attack all arguments for an existing element, and at the same time create a new argument that contains a different version of the element being attacked.

We present two algorithms for replacing elements: the first one replaces the decomposition type (CQ5c, CQ10c), and the second one replaces the name.

\begin{algorithm}[h]
  \caption{CQ5c: Is the decomposition type of element $ie_i$ correct? No, it should be $X$ }\label{alg:replace1}
  \begin{algorithmic}[1]
    \Procedure{$CQ5c$}{$ie_i, X$}
    \State $A \leftarrow ((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$\label{alg:replace1:arg}
    \State $IEArgs\leftarrow \{(ie_i,ie_i.type)\in Arg\}$\label{alg:replace1:ieargs}
    \For{$A_i\in IEArgs$}\label{alg:replace1:for1}
      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:replace1:att}
    \EndFor
    \For{$\{(A_i,A_j)\in Att\mid A_j\in IEArgs\}$}\label{alg:replace1:for2}
      \State $Att\leftarrow (A_i,A)$
    \EndFor
    \For{$\{(A_i,A_j)\in Att\mid A_i\in IEArgs\}$}\label{alg:replace1:for3}
      \State $Att\leftarrow (A,A_j)$
    \EndFor
    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:replace1:addarg}
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}.

Let us go through this algorithm step by step. On line~\ref{alg:replace1:arg}, a new argument $((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$ is created. Recall from Def.~\ref{def:notation} that we may refer to elements of an IE with the ``.'' notation. Thus, this argument is an argument for the IE that was input to the algorithm, except that the decomposition type is set to $X$. On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE. Then, in the first for loop on line~\ref{alg:replace1:for1}, we add attack links from the argument that has just been created to all existing arguments for the IE. The next two for loops on respectively lines~\ref{alg:replace1:for2} and~\ref{alg:replace1:for3} ensure that all attack links that existing from and to the previous versions of the IE are also carried over to the new argument $A$. Finally, one line~\ref{alg:replace1:addarg}, the new argument is added to the set of arguments.

Let us explain this with an example. \todo{Marc}{Marc}{Add example}

The other \emph{REPLACE} algorithms are very similar to Algorithm~\ref{alg:replace1}. It can be used directly for CQ10c, but for CQ12 we should make a small modification. Instead of replacing the decomposition type of the IE, we should replace its name. Since this is a very minor modification we have omitted it here.

### 2017-10-23 methodology.tex
finished all algorithms

The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}. There are two main activities, depicted in grey, namely "goal modeling" and "argumentation". These are two separate activities that are being done in parallel.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/methodology}
\caption{The RationalGRL Methodology}
\label{fig:rationalgrl-methodology}
\end{figure}

>>>>

In the previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models. In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology\footnote{The methodology presented here was first presented at the 2017 iStar workshop \cite{iSTar}.} (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).

\subsection{RationalGRL Methodology}
\label{sect:methodology}

We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model. %Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders. The rest of the steps are as follows:

\begin{figure*}[ht]
\centering
\includegraphics{img/methodology.pdf}
\caption{The RationalGRL Methodology}
\label{fig:rationalgrl-methodology}
\end{figure*}

\textbf{(1) Instantiate Argument Schemes (AS)} -- First, we start with our list of argument schemes (Table~\ref{table:argument-schemes}). Whilst discussing the requirements, we then select schemes from the list and instantiate them to form arguments for GRL model elements. In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}, see example 1 in Section~\ref{sect:overview:examples}). Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e. an instantiated scheme), so it is possible to instantiate argument schemes based on an existing GRL model.

\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we can start asking the relevant critical questions. Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.

\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of three operations are performed on the GRL model: \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}. Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model (see examples in Section~\ref{sect:overview:examples}). After these modifications, we can keep on asking critical questions (e.g. about elements that were introduced by previously answering a critical question) until we are satisfied with our model.

\textbf{(4) Modify GRL Models} -- In this step, we modify the regular GRL model based on the RationalGRL model of step (3). That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one. This results in a new, modified GRL model, which can be used as the basis for another cycle of the methodology.

We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model. In the next section, we will give an example of how our tool can be used together with the methodology to build a GRL model.

\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Example 6: Methodology example}

In this section, we present a small example for the RationalGRL methodology. This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}. We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system. %These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}. %TODO again?
%We provide several more examples in Section~\ref{sect:examples}.

As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \emph{Simulate Traffic} using the RationalGRL methodology. This proceeds in the following way:
\begin{itemize}
\item
First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \emph{Simulate Traffic}. 
\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}). They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
\item As a result, two argument schemes are created, namely:
\begin{itemize}
\item Actor \emph{Traffic Simulator} has task \emph{Dynamic Simulation}
\item Actor \emph{Traffic Simulator} has task \emph{Static Simulation}
\end{itemize}
\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
\begin{itemize}
\item
CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''. Thus, the corresponding GRL intentional element is disabled.
\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
\end{itemize}
\end{itemize}

An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}. %TODO You need to update goal "simulate" in this figure to "Simulate Traffic".  Also, I don't think we should really call it the rational GRL model. it is an excerpts of the model. 
\fi
